---
title: "03_disturbance_and_succession"
author: "Adriana Uscanga"
date: "7/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and files:

```{r}
library(tidyverse)
library(ggplot2)
library(vegan)
library(factoextra)
library(gridExtra)
library(NbClust)
library(goeveg)

load("output/sites_cf2.RData")
load("output/plots_cf2.RData")
load("output/trees_cf2.RData")
```

# Forest Disturbance and Succession

## Forest Disturbance

Read and order FI forest disturbance dataset.

```{r}
disturbance <- load("output/disturbance_sno.RData")

disturbance_cf <- sites_cf %>%
  select(site) %>%
  left_join(disturbance_sno, by = "site") %>%
  select(-c(on_soil, on_water))

disturbance_cf
rm(disturbance_sno)
  
```

Analyze the types of disturbance recorded

```{r}

disturbance_cf %>%
  group_by(on_veg, cause) %>%
  summarize(cause = unique(cause),
            cause_no = n())

disturbance_cf %>%
  group_by(cause, on_veg) %>%
  summarize(no_sites = n()) %>%
  #filter(on_veg == "Mayor") %>%
  ggplot(aes(x= cause, y= no_sites, fill= on_veg)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  theme(axis.text.x = element_text(angle=60, vjust = 1, hjust = 1))

disturbance_cf %>%
  group_by(cause, notes) %>%
  summarize(notes = unique(notes))


#Disturbances related to agriculture and cattle ranches:
#Aprovechamientos forestales
#Incendios
#Pastoreo
#Uso del suelo diferente al forestal

disturbance_cf %>%
  group_by(site) %>%
  summarize(sites_list = unique(site))

disturbance_cf %>%
  filter(is.na(cause)) #14
```


```{r}

agriculture_grazing <- disturbance_cf %>%
  mutate(cause = ifelse(is.na(cause), "unknown", cause)) %>% #Keep sites that were not recorded on disturbance data set
  filter(cause == "Uso del suelo diferente al forestal" | #LULCC
           cause == "Aprovechamientos forestales" | #Logging
           cause == "Incendios" | #Fire -- swidden ag
           cause == "Pastoreo" | # Cattle grazing
           cause == "unknown") %>% #disturbance not reported
  mutate(cause = ifelse(cause == "Uso del suelo diferente al forestal", "lulcc",
                        ifelse(cause =="Aprovechamientos forestales", "logging",
                               ifelse(cause == "Incendios", "fire",
                                      ifelse(cause == "Pastoreo", "grazing", "unknown"))))) %>%
  mutate(cause = ifelse(cause == "logging" & 
                          notes == "Tala para sembrar café." |
                          notes == "Roza-Tumba y quema." |
                          notes == "Roza tumba quema cambio a agricultura temporal.", "agriculture", cause)) %>%
  mutate(cause = ifelse(cause == "fire" &
                          notes == "Quemas agrícolas." |
                          notes == "Rosa, tumba y quema" |
                          notes == "Quema agrícola.", "agriculture", cause)) %>%
  filter(cause == "lulcc" |
           cause == "grazing" |
           cause == "agriculture" |
           is.na(cause)) %>%
  mutate(cause = ifelse(cause == "lulcc" &
                           notes == "Potrero" |
                           notes == "Forestal con ganadería" |
                           notes == "Forestal con ganadería extensiva" |
                           notes == "De forestal a potreros" |
                           notes == "Forestal con ganaderia extensiva." |
                           notes == "Forestal con ganadería extensiva y cultivo de maíz", "grazing", cause)) %>%
  mutate(cause = ifelse(cause == "lulcc", "agriculture", cause)) %>%
  mutate(disturbance_type = as.character(cause),
         disturbance_severity = as.character(on_veg),
         disturbance_notes = as.character(notes)) %>%
  select(-state, -municipality, -cause, -on_veg, -notes) %>%
  mutate(disturbance_type = ifelse(is.na(disturbance_type), "unknown", disturbance_type)) %>%
  mutate(disturbance_severity = ifelse(disturbance_severity == "Mayor", 4,
                                       ifelse(disturbance_severity == "Mediana", 3,
                                              ifelse(disturbance_severity == "Menor", 2,
                                                     ifelse(disturbance_severity == "No perceptible", 1, NA))))) %>%
  mutate(disturbance_severity = as.numeric(disturbance_severity))
  
  
#Notes (causes in logging related to agriculture):
#Tala para sembrar café.
#Roza-Tumba y quema. 
#Roza tumba quema cambio a agricultura temporal.

#Notes in fire:
#Quemas agrícolas.
#Rosa, tumba y quema
#Quema agrícola.
#Notes in lulcc that are grazing:  
#Potrero
#Forestal con ganadería
#Forestal con ganadería extensiva
#De forestal a potreros
#Forestal con ganaderia extensiva.
#Forestal con ganadería extensiva y cultivo de maíz

agriculture_grazing %>%
  group_by(site) %>%
  summarize(sites = unique(site))

#51 sites have either agricultural or grazing activities or both


```

51 sites have either agricultural or grazing activities and 14 are missing from the disturbance data set

Join to sites

```{r}

sites_cf <- agriculture_grazing %>%
  select(-disturbance_notes) %>%
  group_by(site) %>%
  summarize(disturbance_type = unique(disturbance_type),
            disturbance_severity = mean(disturbance_severity)) %>%
  mutate(disturbance_severity = ifelse(disturbance_type == "unknown", -1, disturbance_severity )) %>%
  pivot_wider(names_from = disturbance_type, values_from = disturbance_severity, values_fill = 0) %>%
  right_join(sites_cf, by = "site") %>%
  mutate(agriculture = ifelse(is.na(agriculture), 0, agriculture),
         unknown = ifelse(is.na(unknown), 0, unknown),
         grazing = ifelse(is.na(grazing), 0, grazing)) %>%
  mutate(unknown = ifelse(unknown == -1, NA, unknown)) %>%
  mutate(agriculture = ifelse(is.na(unknown), NA, agriculture),
         grazing = ifelse(is.na(unknown), NA, grazing)) %>%
  select(-unknown) %>%
  relocate(c("agriculture", "grazing"), .after = "sp_richness_site_sd")

```

## Forest Succession

### Non-hierarchical cluster analysis and PCA

Select structure attributes and scale dataset so that variables in different units can be compared

```{r}

# Plots

# tree_no_ave and av_treedensity_site are completly correlated.
# Remove tree_no_ave to avoid correlation in further analyses.

plots_str <- plots_cf %>%
  select(plot_id,
         tree_density,
         loreys_height,
         basal_area_ha,
         agb_plot_ha)

plot(plots_str)

# Scale data so that variables in different units can be compared
plots_str_scaled <- scale(plots_str[,-c(1,5)])
row.names(plots_str_scaled) <- plots_str$plot_id

# Repeat with a subset of 160 plots (for the sites that have 4 plots each)


plots_str <- plots_cf %>%
  select(plot_id,
         tree_density,
         loreys_height,
         basal_area_ha,
         agb_plot_ha)

plot(plots_str)

# Scale data so that variables in different units can be compared
plots_str_scaled <- scale(plots_str[,-c(1,5)])
row.names(plots_str_scaled) <- plots_str$plot_id

```

Compute K-means with several number of clusters and compare plots

```{r}

# Plots
k2_structure_plots <- kmeans(plots_str_scaled, centers = 2, nstart = 25)
k3_structure_plots <- kmeans(plots_str_scaled, centers = 3, nstart = 25)
k4_structure_plots <- kmeans(plots_str_scaled, centers = 4, nstart = 25)
k5_structure_plots <- kmeans(plots_str_scaled, centers = 5, nstart = 25)

# Compare plots:
p1_str_plots <- fviz_cluster(k2_structure_plots, geom = "point", data = plots_str_scaled) + ggtitle("k = 2")
p2_str_plots <- fviz_cluster(k3_structure_plots, geom = "point",  data = plots_str_scaled) + ggtitle("k = 3")
p3_str_plots <- fviz_cluster(k4_structure_plots, geom = "point",  data = plots_str_scaled) + ggtitle("k = 4")
p4_str_plots <- fviz_cluster(k5_structure_plots, geom = "point",  data = plots_str_scaled) + ggtitle("k = 5")
grid.arrange(p1_str_plots, p2_str_plots, p3_str_plots, p4_str_plots, nrow = 2)

```

Data can be classified in many different number of clusters.
To define the best number of clusters, we're using NbClust.

```{r}
# Plots

str_clusters_plots <- NbClust((plots_str_scaled^2), distance = "euclidean", min.nc = 2, max.nc = 10, method = "ward.D2", index = "all")


```

Most indices proposed 3 as the best number of clusters (for both sites and plots).
Now I can classify data in three classes. Also, I can visualize these three classes in a PCA to see how structure attributes defined them.

```{r}
# Plots

# Save a vector of 3 clusters:
str_k3_plots <- k3_structure_plots$cluster
as.numeric(str_k3_plots)

#pca with structural data

pca_str_plots <- rda(plots_str_scaled, scale = F)
summary(pca_str_plots) # 
plot(pca_str_plots)

ordiplot(pca_str_plots, display = 'sites', type = 'n')
points(pca_str_plots, col = str_k3_plots)
biplot(pca_str_plots, display = 'species')

ordiplot(pca_str_plots, display = 'sites', type = 'text')
```

Add cluster class plots_cf dataset.

```{r}

# Plots

str_k3_plots <- as_tibble(str_k3_plots, rownames = "plot_id")

str_k3_plots <- str_k3_plots %>%
  mutate(str_class_plots_4p = as.character(value)) %>%
  select(-value)

plots_cf <- plots_cf %>%
  left_join(str_k3_plots, by = "plot_id")


```

### Strucutre attributes across clusters 

Let's see how structure attributes change in these three clusters.

```{r}

# Plots

plots_cf %>%
  ggplot(aes(x = tree_density, group = str_class_plots, fill = str_class_plots, color = str_class_plots)) +
  geom_density(alpha= 0.5) +
  scale_fill_manual(values = c("tomato", "green3", "dodgerblue2")) +
  scale_color_manual(values = c("tomato", "green3", "dodgerblue2")) +
  ggtitle("Tree density")

plots_cf %>%
  ggplot(aes(x = loreys_height, group = str_class_plots, fill = str_class_plots, color = str_class_plots)) +
  geom_density(alpha= 0.5) +
  scale_fill_manual(values = c("tomato", "green3", "dodgerblue2")) +
  scale_color_manual(values = c("tomato", "green3", "dodgerblue2")) +
  ggtitle("Tree height")

plots_cf %>%
  ggplot(aes(x = basal_area_ha, group = str_class_plots, fill = str_class_plots, color = str_class_plots)) +
  geom_density(alpha= 0.5) +
  scale_fill_manual(values = c("tomato", "green3", "dodgerblue2")) +
  scale_color_manual(values = c("tomato", "green3", "dodgerblue2")) +
  ggtitle("Basal area")

#plots_cf %>%
#  ggplot(aes(x = agb_plot_ha, group = str_class_plots, fill = str_class_plots)) +
#  geom_density(alpha= 0.5) +
#  ggtitle("Biomass")


#################This code below will probably go to the 'extra code bucket' ###############

# Plots :4p

# plots_cf_4p %>%
#   ggplot(aes(x = tree_density, group = str_class_plots_4p, fill = str_class_plots_4p)) +
#   geom_density(alpha= 0.5) +
#   scale_fill_manual(values = c("tomato", "green3", "dodgerblue2")) +
#   ggtitle("Tree density")
# 
# plots_cf_4p %>%
#   ggplot(aes(x = loreys_height, group = str_class_plots_4p, fill = str_class_plots_4p)) +
#   geom_density(alpha= 0.5) +
#   scale_fill_manual(values = c("tomato", "green3", "dodgerblue2")) +
#   ggtitle("Tree height")
# 
# plots_cf_4p %>%
#   ggplot(aes(x = basal_area_ha, group = str_class_plots_4p, fill = str_class_plots_4p)) +
#   geom_density(alpha= 0.5) +
#   scale_fill_manual(values = c("tomato", "green3", "dodgerblue2")) +
#   ggtitle("Basal area")

```

```{r}

# Plots

plots_cf %>%
  ggplot(aes(y = tree_density, color = str_class_plots)) +
  geom_boxplot()

plots_cf %>%
  ggplot(aes(y = loreys_height, color = str_class_plots)) +
  geom_boxplot()

plots_cf %>%
  ggplot(aes(y = basal_area_ha, color = str_class_plots)) +
  geom_boxplot()

plots_cf %>%
  ggplot(aes(y = agb_plot_ha, color = str_class_plots)) +
  geom_boxplot()

plots_cf %>%
  ggplot(aes(y = shannon, color = str_class_plots)) +
  geom_boxplot()

plots_cf %>%
  ggplot(aes(y = simpson, color = str_class_plots)) +
  geom_boxplot()
```

* Note: there's a plot with more than 600 Mg/ha of biomass, which sounds like too much. Perhaps there's a mistake in the data. *

Structure attributes are clearly different among clusters. It seems that plots split in those with very few trees, many trees of medium size, and plots with very big trees with medium tree density. This broadly coincides with succession stages: very young, young-medium, and mature forests. 

### Landscape composition: mosaic description

```{r}

mosaic <- plots_cf %>%
  mutate(str_class_plots = as.numeric(str_class_plots),
         site = as.character(site)) %>%
  group_by(site) %>%
  # normalized mosaic = value - min value / max value - min value
  # mosaic goes from 1 (all mature forest) to 0 (all agriculture/very young forest)
  summarize(mosaic4 = ifelse(n()==4, (sum(str_class_plots)-4)/(12-4), NA),
            mosaic3 = ifelse(n()==3, (sum(str_class_plots)-3)/(9-3), NA),
            mosaic2 = ifelse(n()==2, (sum(str_class_plots)-2)/(6-2), NA),
            mosaic1 = ifelse(n()==1, (sum(str_class_plots)-1)/(3-1), NA),
            mosaic = ifelse(n()==4, mosaic4,
                            ifelse(n()==3, mosaic3,
                                   ifelse(n()==2, mosaic2, mosaic1)))) %>%
  arrange(mosaic)

sites_cf <- sites_cf %>%
  left_join(mosaic, by = "site") %>%
  select(-c(mosaic1, mosaic2, mosaic3, mosaic4))

sites_cf %>%
  filter(plot_no == 4) %>%
  ggplot(aes(x= mosaic, y = site)) +
  geom_point()

sites_cf %>%
  filter(plot_no == 4,
         mosaic == 0.5)

sites_cf %>%
  filter(plot_no >3) %>%
  group_by(mosaic) %>%
  summarize(agb_mosaic = mean(av_agb_site), av_agb_site = as.numeric(av_agb_site)) %>%
  ggplot(aes(x = mosaic, y = av_agb_site)) +
  geom_point() +
  geom_point(aes(x= mosaic, y = agb_mosaic), color = "red")
  
save(sites_cf, file = "output/sites_cf3.RData")
save(plots_cf, file = "output/plots_cf3.RData")
save(trees_cf, file = "output/trees_cf3.RData")

#################This code below will probably go to the 'extra code bucket' ###############

# Mosaic sites with 4p

# mosaic_4p <- plots_cf_4p %>%
#   mutate(str_class_plots_4p = as.numeric(str_class_plots_4p),
#          site = as.character(site)) %>%
#   group_by(site) %>%
#   # normalized mosaic = value - min value / max value - min value
#   # mosaic goes from 1 (all mature forest) to 0 (all agriculture/very young forest)
#   summarize(mosaic4 = ifelse(n()==4, (sum(str_class_plots_4p)-4)/(12-4), NA),
#             mosaic3 = ifelse(n()==3, (sum(str_class_plots_4p)-3)/(9-3), NA),
#             mosaic2 = ifelse(n()==2, (sum(str_class_plots_4p)-2)/(6-2), NA),
#             mosaic1 = ifelse(n()==1, (sum(str_class_plots_4p)-1)/(3-1), NA),
#             mosaic_4p = ifelse(n()==4, mosaic4,
#                             ifelse(n()==3, mosaic3,
#                                    ifelse(n()==2, mosaic2, mosaic1)))) %>%
#   arrange(mosaic_4p)
# 
# sites_cf_4p <- sites_cf_4p %>%
#   left_join(mosaic_4p, by = "site") %>%
#   select(-c(mosaic1, mosaic2, mosaic3, mosaic4))
# 
# plot(sites_cf_4p$mosaic, sites_cf_4p$mosaic_4p)
# 
# plot(sites_cf_4p$mosaic_4p,sites_cf_4p$av_agb_site)
# plot(sites_cf_4p$mosaic_4p, sites_cf_4p$simpson_av_site)


```

################################################################################
#######################     EXTRA CODE BUCKET    ###############################
################################################################################

### Species composition and diversity across classes

First, add classes info to trees_cf dataset

```{r}

load("output/sites_cf3.RData")
load("output/plots_cf3.RData")
load("output/trees_cf3.RData")


# Make site column in tree_cf dataset

trees_cf <- trees_cf %>%
  separate(plot_id, c("site", "plot"), sep = "_", remove = FALSE)

# Join datasets (sites)

trees_cf <- sites_cf %>%
  select(site, mosaic) %>%
  right_join(trees_cf, by = "site")

# Join datasets (plots)

trees_cf <- plots_cf %>%
  select(plot_id, str_class_plots) %>%
  right_join(trees_cf, by = "plot_id")
```

There are two possible paths now:
1. Make a RDA (linear data) or CCA (unimodal data) with structure attributes and sp composition--didn't work too good. Spp composition adds too much variation. Many species are in very few plots. It's difficult to get general patterns.

2. Classify data using the 3 structural classes (from kmeans analysis) and assess sp abundance and richness to see if there's a logical relationship.


First, make a matrix of sp abundace

```{r}

# Sites

sp_site <- trees_cf %>% # make a table of community data for estimating diversity
  filter(life_form == "Arbol" |
           is.na(life_form)) %>%
  separate(plot_id, c("site", "plot"), sep = "_", remove = F) %>%
  dplyr::group_by(site, species) %>%
  dplyr::summarize(abundance = sum(n())) %>%
  filter(!is.na(species)) %>%
  spread(species, abundance, fill = 0) %>%
  column_to_rownames(var = "site")

#save(sp_site, file = "output/sp_site.RData") #Sites by Species Matrix

sp_site2 <- trees_cf %>% # make a table of community data for estimating diversity
  filter(life_form == "Arbol" |
           is.na(life_form)) %>%
  separate(plot_id, c("site", "plot"), sep = "_", remove = F) %>%
  dplyr::group_by(site, species) %>%
  dplyr::summarize(abundance = sum(n())) %>%
  filter(!is.na(species))

# Plots

sp_site_plots <- trees_cf %>% # make a table of community data for estimating diversity
  filter(life_form == "Arbol" |
           is.na(life_form)) %>%
  dplyr::group_by(plot_id, species) %>%
  dplyr::summarize(abundance = sum(n())) %>%
  filter(!is.na(species)) %>%
  spread(species, abundance, fill = 0) %>%
  column_to_rownames(var = "plot_id")

#save(sp_site, file = "output/sp_site.RData") #Sites by Species Matrix

sp_site2_plots <- trees_cf %>% # make a table of community data for estimating diversity
  filter(life_form == "Arbol" |
           is.na(life_form)) %>%
  dplyr::group_by(plot_id, species) %>%
  dplyr::summarize(abundance = sum(n())) %>%
  filter(!is.na(species))
```

Now let's analyze the abundance of species in the three classes (site and plot levels).

```{r}

# Sites

community_matureforest_4p <- sites_cf %>%
  filter(plot_no == 4) %>%
  filter(mosaic > 0.75) %>%
  select(site) %>%
  left_join(as_tibble(sp_site, rownames = "site"), by = "site") 

community_matureforest_4p <- as.data.frame(community_matureforest_4p)
row.names(community_matureforest_4p) <- community_matureforest_4p$site
community_matureforest_4p <- community_matureforest_4p[,-1]

community_secandmature_4p <- sites_cf %>%
  filter(plot_no == 4) %>%
  filter(mosaic > 0.5, mosaic < 0.76) %>%
  select(site) %>%
  left_join(as_tibble(sp_site, rownames = "site"), by = "site") 

community_secandmature_4p <- as.data.frame(community_secandmature_4p)
row.names(community_secandmature_4p) <- community_secandmature_4p$site
community_secandmature_4p <- community_secandmature_4p[,-1]


community_class2 <- sites_cf %>%
  filter(plot_no > 1) %>%
  filter(str_class == "2") %>%
  select(site) %>%
  left_join(as_tibble(sp_site, rownames = "site"), by = "site")

community_class2 <- as.data.frame(community_class2)
row.names(community_class2) <- community_class2$site
community_class2 <- community_class2[,-1]

community_class3 <- sites_cf %>%
  filter(plot_no > 1) %>%
  filter(str_class == "3") %>%
  select(site) %>%
  left_join(as_tibble(sp_site, rownames = "site"), by = "site")

community_class3 <- as.data.frame(community_class3)
row.names(community_class3) <- community_class3$site
community_class3 <- community_class3[,-1]

# Plots

community_class1_plots <- plots_cf %>%
  filter(str_class_plots == "1") %>%
  select(plot_id) %>%
  left_join(as_tibble(sp_site_plots, rownames = "plot_id"), by = "plot_id") 

community_class1_plots <- as.data.frame(community_class1_plots)
row.names(community_class1_plots) <- community_class1_plots$plot_id
community_class1_plots <- community_class1_plots[,-1]
  
community_class2_plots <- plots_cf %>%
  filter(str_class_plots == "2") %>%
  select(plot_id) %>%
  left_join(as_tibble(sp_site_plots, rownames = "plot_id"), by = "plot_id")

community_class2_plots <- as.data.frame(community_class2_plots)
row.names(community_class2_plots) <- community_class2_plots$plot_id
community_class2_plots <- community_class2_plots[,-1]

community_class3_plots <- plots_cf %>%
  filter(str_class_plots == "3") %>%
  select(plot_id) %>%
  left_join(as_tibble(sp_site_plots, rownames = "plot_id"), by = "plot_id")

community_class3_plots <- as.data.frame(community_class3_plots)
row.names(community_class3_plots) <- community_class3_plots$plot_id
community_class3_plots <- community_class3_plots[,-1]
```

Remove spp with abundance = 0 in each class

```{r}

# Sites

community_matureforest_4p <- community_matureforest_4p[, which(colSums(community_matureforest_4p) != 0)]
community_secandmature_4p <- community_secandmature_4p[, which(colSums(community_secandmature_4p) != 0)]
community_class2 <- community_class2[, which(colSums(community_class2) != 0)]
community_class3 <- community_class3[, which(colSums(community_class3) != 0)]

# Plots

community_class1_plots <- na.omit(community_class1_plots)
community_class1_plots <- community_class1_plots[, which(colSums(community_class1_plots) != 0)]

community_class2_plots <- na.omit(community_class2_plots)
community_class2_plots <- community_class2_plots[, which(colSums(community_class2_plots) != 0)]

community_class3_plots <- na.omit(community_class3_plots)
community_class3_plots <- community_class3_plots[,which(colSums(community_class3_plots) != 0)]

```

Use racurve function to plot rank-abundance curves

```{r}
racurve1 <- racurve(community_matureforest_4p, main = "Mature Forest", nlab = 8)
racurve2 <- racurve(community_secandmature_4p, main = "Secondary and Mature Forest", nlab = 8)
racurve2 <- racurve(community_class2, main = "Class 2", nlab = 5)
racurve3 <- racurve(community_class3, main = "Class 3", nlab = 5)

racurve(community_class2[1,], nlab = 5)
racurve(community_class2[2,], nlab = 5)
racurve(community_class2[3,], nlab = 5)
racurve(community_class2[4,], nlab = 5)
racurve(community_class2[5,], nlab = 5)
racurve(community_class2[6,], nlab = 5)
racurve(community_class2[7,], nlab = 5)

racurve(community_class3["70980",], nlab = 5)

# Plots
racurve1_plots <- racurve(community_class1_plots, main = "Class 1", nlab = 6)
racurve2_plots <- racurve(community_class2_plots, main = "Class 2", nlab = 6)
racurve3_plots <- racurve(community_class3_plots, main = "Class 3", nlab = 6)

```

```{r}
racurve3_plots
```

Relative abundance:

```{r}
racurve1$rel.abund
racurve2$rel.abund
racurve3$rel.abund

racurve1_plots$rel.abund
racurve2_plots$rel.abund
racurve3_plots$rel.abund
```

Let's take a closer look at each class
```{r}
sites_cf %>%
  filter(!is.na(str_class)) %>%
  filter(str_class == "1")

sites_cf %>%
  filter(!is.na(str_class)) %>%
  ggplot(aes(x = str_class, y = altitude)) +
  geom_point()

structure_forest <- sites_cf %>%
  filter(str_class == "2" | str_class == "3") %>%
  select(site,
         av_basalarea_site,
         av_loreys_site,
         av_treedensity_site)

# Scale data so that variables in different units can be compared
structure_forest_scaled <- scale(structure_forest[,-c(1)])
row.names(structure_forest_scaled) <- structure_forest$site

k2_structure_forest <- kmeans(structure_forest_scaled, centers = 2, nstart = 25)
k3_structure_forest <- kmeans(structure_forest_scaled, centers = 3, nstart = 25)
k4_structure_forest <- kmeans(structure_forest_scaled, centers = 4, nstart = 25)
k5_structure_forest <- kmeans(structure_forest_scaled, centers = 5, nstart = 25)

# Compare plots:
p1_str_f <- fviz_cluster(k2_structure_forest, geom = "point", data = structure_forest_scaled) + ggtitle("k = 2")
p2_str_f <- fviz_cluster(k3_structure_forest, geom = "point",  data = structure_forest_scaled) + ggtitle("k = 3")
p3_str_f <- fviz_cluster(k4_structure_forest, geom = "point",  data = structure_forest_scaled) + ggtitle("k = 4")
p4_str_f <- fviz_cluster(k5_structure_forest, geom = "point",  data = structure_forest_scaled) + ggtitle("k = 5")
grid.arrange(p1_str, p2_str, p3_str, p4_str, nrow = 2)


str_clusters_f <- NbClust((structure_forest_scaled^2), distance = "euclidean", min.nc = 2, max.nc = 10, method = "ward.D2", index = "all")

# Save a vector of 3 clusters:
str_k3_forest <- k3_structure_forest$cluster
as.numeric(str_k3_forest)

#pca with structural data

pca_str_f <- rda(structure_forest_scaled, scale = F)
summary(pca_str_f) 
plot(pca_str_f)

ordiplot(pca_str_f, display = 'sites', type = 'n')
points(pca_str_f, col = str_k3_forest)
biplot(pca_str_f, display = 'species')


```

Let's see if community data follows a similar classification than structural data

```{r}
# Calculate a distance matrix

log_sp_site <- log1p(sp_site)
dm_euclidean <- vegdist(log_sp_site, method = "euclidean")
cluster_ward <- hclust(dm_euclidean, method = "ward.D")
plot(cluster_ward)

library(dendextend)
dend_cluster_ward <- as.dendrogram(cluster_ward)
labels_colors(dend_cluster_ward)
labels_colors(dend_cluster_ward) <- as.numeric(sites_cf$str_class)
dend_cluster_ward <- set(dend_cluster_ward, "labels_cex", 0.5)
get_leaves_nodePar(dend_cluster_ward)[[1]]
plot(dend_cluster_ward)

```

