---
title: "03_disturbance_and_succession"
author: "Adriana Uscanga"
date: "7/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and files:

```{r}
library(tidyverse)
library(ggplot2)
library(vegan)
library(factoextra)
library(gridExtra)
library(NbClust)
library(goeveg)

load("output/sites_cf2.RData")
load("output/plots_cf2.RData")
load("output/trees_cf2.RData")
```

# Forest Disturbance and Succession

## Forest Disturbance

Read and order FI forest disturbance dataset.

```{r}
disturbance <- load("output/disturbance_sno.RData")

disturbance_cf <- sites_cf %>%
  select(site) %>%
  left_join(disturbance_sno, by = "site") %>%
  select(-c(on_soil, on_water))

disturbance_cf
  
```
Analyze the types of disturbance recorded

```{r}

disturbance_cf %>%
  group_by(on_veg, cause) %>%
  summarize(cause = unique(cause),
            cause_no = n())

disturbance_cf %>%
  group_by(cause, on_veg) %>%
  summarize(no_sites = n()) %>%
  #filter(on_veg == "Mayor") %>%
  ggplot(aes(x= cause, y= no_sites, fill= on_veg)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  theme(axis.text.x = element_text(angle=60, vjust = 1, hjust = 1))

disturbance_cf %>%
  group_by(cause, notes) %>%
  summarize(notes = unique(notes))

dist_sev <- disturbance_cf %>%
  group_by(site) %>%
  summarize(disturbance_site = unique(on_veg)) %>%
  mutate(disturbance_severity = ifelse(disturbance_site == "Mayor", 3,
                                       ifelse(disturbance_site == "Mediana", 2,
                                              ifelse(disturbance_site == "Menor", 1,
                                                     ifelse(disturbance_site == "No perceptible", 0, NA))))) %>%
  filter(!is.na(disturbance_severity)) %>%
  group_by(site) %>%
  summarize(no_disturbance = n(),
            dist_sev = ((sum(disturbance_severity)/no_disturbance)* no_disturbance))
```

Summarize disturbance info by site
NDVI time series

## Forest Succession

### Non-hierarchical cluster analysis and PCA

Select structure attributes and scale dataset so that variables in different units can be compared

```{r}

#Sites

sites_str <- sites_cf %>%
  filter(plot_no > 1) %>%
  select(site,
         tree_no_ave,
         av_basalarea_site,
         av_loreys_site,
         av_treeheight_site,
         av_treedensity_site)

sites_str
plot(sites_str)

# tree_no_ave and av_treedensity_site are completly correlated.
# Remove tree_no_ave to avoid correlation in further analyses.

sites_str <- sites_cf %>%
  filter(plot_no > 1) %>%
  select(site,
         av_basalarea_site,
         av_loreys_site,
         av_treedensity_site)

# Scale data so that variables in different units can be compared
sites_str_scaled <- scale(sites_str[,-(1)])
row.names(sites_str_scaled) <- sites_str$site

# Plots

plots_str <- plots_cf %>%
  select(plot_id,
         tree_density,
         loreys_height,
         basal_area_ha,
         agb_plot_ha)

plot(plots_str)

plots_str_scaled <- scale(plots_str[,-c(1,5)])
row.names(plots_str_scaled) <- plots_str$plot_id

```


Compute K-means with several number of clusters and compare plots

```{r}

# Sites

k2_structure <- kmeans(sites_str_scaled, centers = 2, nstart = 25)
k3_structure <- kmeans(sites_str_scaled, centers = 3, nstart = 25)
k4_structure <- kmeans(sites_str_scaled, centers = 4, nstart = 25)
k5_structure <- kmeans(sites_str_scaled, centers = 5, nstart = 25)

# Compare plots:
p1_str <- fviz_cluster(k2_structure, geom = "point", data = sites_str_scaled) + ggtitle("k = 2")
p2_str <- fviz_cluster(k3_structure, geom = "point",  data = sites_str_scaled) + ggtitle("k = 3")
p3_str <- fviz_cluster(k4_structure, geom = "point",  data = sites_str_scaled) + ggtitle("k = 4")
p4_str <- fviz_cluster(k5_structure, geom = "point",  data = sites_str_scaled) + ggtitle("k = 5")
grid.arrange(p1_str, p2_str, p3_str, p4_str, nrow = 2)

# Plots
k2_structure_plots <- kmeans(plots_str_scaled, centers = 2, nstart = 25)
k3_structure_plots <- kmeans(plots_str_scaled, centers = 3, nstart = 25)
k4_structure_plots <- kmeans(plots_str_scaled, centers = 4, nstart = 25)
k5_structure_plots <- kmeans(plots_str_scaled, centers = 5, nstart = 25)

# Compare plots:
p1_str_plots <- fviz_cluster(k2_structure_plots, geom = "point", data = plots_str_scaled) + ggtitle("k = 2")
p2_str_plots <- fviz_cluster(k3_structure_plots, geom = "point",  data = plots_str_scaled) + ggtitle("k = 3")
p3_str_plots <- fviz_cluster(k4_structure_plots, geom = "point",  data = plots_str_scaled) + ggtitle("k = 4")
p4_str_plots <- fviz_cluster(k5_structure_plots, geom = "point",  data = plots_str_scaled) + ggtitle("k = 5")
grid.arrange(p1_str_plots, p2_str_plots, p3_str_plots, p4_str_plots, nrow = 2)

```
Data can be classified in many different number of clusters.
To define the best number of clusters, we're using NbClust.

```{r}
#Sites

str_clusters <- NbClust((sites_str_scaled^2), distance = "euclidean", min.nc = 2, max.nc = 10, method = "ward.D2", index = "all")

# Plots

str_clusters_plots <- NbClust((plots_str_scaled^2), distance = "euclidean", min.nc = 2, max.nc = 10, method = "ward.D2", index = "all")


```

Most indices proposed 3 as the best number of clusters (for both sites and plots).
Now I can classify data in three classes. Also, I can visualize these three classes in a PCA to see how structure attributes defined them.

```{r}
# Sites

# Save a vector of 3 clusters:
str_k3 <- k3_structure$cluster
as.numeric(str_k3)

#pca with structural data

pca_str <- rda(sites_str_scaled, scale = F)
summary(pca_str) # PC1 explains 67.69% of variation and PC2 25.04%
plot(pca_str)

ordiplot(pca_str, display = 'sites', type = 'n')
points(pca_str, col = str_k3)
biplot(pca_str, display = 'species')

# Plots

# Save a vector of 3 clusters:
str_k3_plots <- k3_structure_plots$cluster
as.numeric(str_k3_plots)

#pca with structural data

pca_str_plots <- rda(plots_str_scaled, scale = F)
summary(pca_str_plots) # 
plot(pca_str_plots)

ordiplot(pca_str_plots, display = 'sites', type = 'n')
points(pca_str_plots, col = str_k3_plots)
biplot(pca_str_plots, display = 'species')

```

Add cluster class to sites_cf and plots_cf dataset.

```{r}

# Sites

str_k3 <- as_tibble(str_k3, rownames = "site")

str_k3 <- str_k3 %>%
  mutate(str_class = as.character(value)) %>%
  select(-value)

sites_cf <- sites_cf %>%
  left_join(str_k3, by = "site")

sites_cf <- sites_cf %>%
  left_join(dist_sev, by = "site")

# Plots

str_k3_plots <- as_tibble(str_k3_plots, rownames = "plot_id")

str_k3_plots <- str_k3_plots %>%
  mutate(str_class_plots = as.character(value)) %>%
  select(-value)

plots_cf <- plots_cf %>%
  left_join(str_k3_plots, by = "plot_id")


```

### Biomass and strucutre attributes across clusters 

Let's see how structure attributes change in these three clusters.

```{r}

# Sites

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(x = tree_no_site, group = str_class, fill = str_class)) +
  geom_density(alpha= 0.5) +
  ggtitle("Tree number")

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(x = av_basalarea_site, group = str_class, fill = str_class)) +
  geom_density(alpha= 0.5) +
  ggtitle("Basal Area")

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(x = av_treeheight_site, group = str_class, fill = str_class)) +
  geom_density(alpha= 0.5) +
  ggtitle("Tree height")

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(x = av_treedensity_site, group = str_class, fill = str_class)) +
  geom_density(alpha= 0.5) +
  ggtitle("Tree density")

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(x = av_agb_site, group = str_class, fill = str_class)) +
  geom_density(alpha= 0.5) +
  ggtitle("Biomass")

# Plots

plots_cf %>%
  ggplot(aes(x = tree_no, group = str_class_plots, fill = str_class_plots)) +
  geom_density(alpha= 0.5) +
  ggtitle("Tree number")

plots_cf %>%
  ggplot(aes(x = loreys_height, group = str_class_plots, fill = str_class_plots)) +
  geom_density(alpha= 0.5) +
  ggtitle("Tree height")

plots_cf %>%
  ggplot(aes(x = basal_area_ha, group = str_class_plots, fill = str_class_plots)) +
  geom_density(alpha= 0.5) +
  ggtitle("Basal area")

plots_cf %>%
  ggplot(aes(x = agb_plot_ha, group = str_class_plots, fill = str_class_plots)) +
  geom_density(alpha= 0.5) +
  ggtitle("Biomass")

```

```{r}

# Sites

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(y = av_treedensity_site, color = str_class)) +
  geom_boxplot()

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(y = tree_no_site, color = str_class)) +
  geom_boxplot()

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(y = av_treeheight_site, color = str_class)) +
  geom_boxplot()

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(y = av_basalarea_site, color = str_class)) +
  geom_boxplot()

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(y = av_agb_site, color = str_class)) +
  geom_boxplot()

sites_cf %>%
  filter(plot_no > 1) %>%
  ggplot(aes(x = str_class, y = dist_sev)) +
  geom_point()

plot(sites_cf$dist_sev)

sites_cf %>%
  filter(plot_no > 1) %>%
  filter(str_class == "2" | str_class == "3") %>%
  ggplot(aes(y = av_agb_site, x = altitude, color = precip, shape = str_class)) +
  geom_point()

# Plots

plots_cf %>%
  ggplot(aes(y = tree_no, color = str_class_plots)) +
  geom_boxplot()

plots_cf %>%
  ggplot(aes(y = loreys_height, color = str_class_plots)) +
  geom_boxplot()

plots_cf %>%
  ggplot(aes(y = basal_area_ha, color = str_class_plots)) +
  geom_boxplot()

plots_cf %>%
  ggplot(aes(y = agb_plot_ha, color = str_class_plots)) +
  geom_boxplot()
```

* Note: there's a plot with more than 600 Mg/ha of biomass, which sounds like too much. Perhaps there's a mistake in the data. *

Structure attributes are clearly different among clusters. It seems that plots split in those with very few trees (small agb), many trees of medium size, and plots with very big trees with medium tree density. This broadly coincides with succession stages: very young, young-medium, and mature forests. 

### Mosaics description

```{r}

mosaic <- plots_cf %>%
  mutate(str_class_plots = as.numeric(str_class_plots),
         site = as.character(site)) %>%
  group_by(site) %>%
  # normalized mosaic = value - min value / max value - min value
  # mosaic goes from 0 (all mature forest) to 1 (all agriculture/very young forest)
  summarize(mosaic4 = ifelse(n()==4, (sum(str_class_plots)-4)/(12-4), NA),
            mosaic3 = ifelse(n()==3, (sum(str_class_plots)-3)/(9-3), NA),
            mosaic2 = ifelse(n()==2, (sum(str_class_plots)-2)/(6-2), NA),
            mosaic1 = ifelse(n()==1, (sum(str_class_plots)-1)/(3-1), NA),
            mosaic = ifelse(n()==4, mosaic4,
                            ifelse(n()==3, mosaic3,
                                   ifelse(n()==2, mosaic2, mosaic1)))) %>%
  arrange(mosaic)

sites_cf <- sites_cf %>%
  left_join(mosaic, by = "site") %>%
  select(-c(mosaic1, mosaic2, mosaic3, mosaic4))

sites_cf %>%
  filter(plot_no == 4) %>%
  ggplot(aes(x= mosaic, y = site)) +
  geom_point()

sites_cf %>%
  filter(plot_no == 4,
         mosaic == 0.5)

sites_cf %>%
  filter(plot_no >3) %>%
  ggplot(aes(x = mosaic, y = av_agb_site)) +
  geom_point()
  
save(sites_cf, file = "output/sites_cf3.RData")
save(plots_cf, file = "output/plots_cf3.RData")
save(trees_cf, file = "output/trees_cf3.RData")

```

###############################################################
**Normality tests, remove from here**
```{r}
# Normality tests
# In summary, tree height is the only normal variable. The others don't have a normal distribution
# when variables are grouped in structural classes, normality increases.

plots_cf %>%
  group_by(str_class_plots) %>%
  shapiro_test(tree_density)

ggqqplot(plots_cf$basal_area_ha)
shapiro_test(plots_cf$basal_area_ha)

sites_cf %>%
  shapiro_test(av_treedensity_site)
ggqqplot(sites_cf$av_agb_site)
ggdensity(sites_cf$av_agb_site)

agblog <- log(sites_cf$av_agb_site)
shapiro.test(agblog)
ggqqplot(agblog)
ggdensity(agblog)

sites_cf %>%
  group_by(str_class) %>%
  shapiro_test(av_agb_site)

plotDist(agblog)

```

### Species composition and diversity across classes

First, add classes info to trees_cf dataset

```{r}

load("output/sites_cf3.RData")
load("output/plots_cf3.RData")
load("output/trees_cf3.RData")


# Make site column in tree_cf dataset

trees_cf <- trees_cf %>%
  separate(plot_id, c("site", "plot"), sep = "_", remove = FALSE)

# Join datasets (sites)

trees_cf <- sites_cf %>%
  select(site, mosaic) %>%
  right_join(trees_cf, by = "site")

# Join datasets (plots)

trees_cf <- plots_cf %>%
  select(plot_id, str_class_plots) %>%
  right_join(trees_cf, by = "plot_id")
```

There are two possible paths now:
1. Make a RDA (linear data) or CCA (unimodal data) with structure attributes and sp composition--didn't work too good. Spp composition adds too much variation. Many species are in very few plots. It's difficult to get general patterns.

2. Classify data using the 3 structural classes (from kmeans analysis) and assess sp abundance and richness to see if there's a logical relationship.


First, make a matrix of sp abundace

```{r}

# Sites

sp_site <- trees_cf %>% # make a table of community data for estimating diversity
  filter(life_form == "Arbol" |
           is.na(life_form)) %>%
  separate(plot_id, c("site", "plot"), sep = "_", remove = F) %>%
  dplyr::group_by(site, species) %>%
  dplyr::summarize(abundance = sum(n())) %>%
  filter(!is.na(species)) %>%
  spread(species, abundance, fill = 0) %>%
  column_to_rownames(var = "site")

#save(sp_site, file = "output/sp_site.RData") #Sites by Species Matrix

sp_site2 <- trees_cf %>% # make a table of community data for estimating diversity
  filter(life_form == "Arbol" |
           is.na(life_form)) %>%
  separate(plot_id, c("site", "plot"), sep = "_", remove = F) %>%
  dplyr::group_by(site, species) %>%
  dplyr::summarize(abundance = sum(n())) %>%
  filter(!is.na(species))

# Plots

sp_site_plots <- trees_cf %>% # make a table of community data for estimating diversity
  filter(life_form == "Arbol" |
           is.na(life_form)) %>%
  dplyr::group_by(plot_id, species) %>%
  dplyr::summarize(abundance = sum(n())) %>%
  filter(!is.na(species)) %>%
  spread(species, abundance, fill = 0) %>%
  column_to_rownames(var = "plot_id")

#save(sp_site, file = "output/sp_site.RData") #Sites by Species Matrix

sp_site2_plots <- trees_cf %>% # make a table of community data for estimating diversity
  filter(life_form == "Arbol" |
           is.na(life_form)) %>%
  dplyr::group_by(plot_id, species) %>%
  dplyr::summarize(abundance = sum(n())) %>%
  filter(!is.na(species))
```

Now let's analyze the abundance of species in the three classes (site and plot levels).

```{r}

# Sites

community_matureforest_4p <- sites_cf %>%
  filter(plot_no == 4) %>%
  filter(mosaic > 0.75) %>%
  select(site) %>%
  left_join(as_tibble(sp_site, rownames = "site"), by = "site") 

community_matureforest_4p <- as.data.frame(community_matureforest_4p)
row.names(community_matureforest_4p) <- community_matureforest_4p$site
community_matureforest_4p <- community_matureforest_4p[,-1]

community_secandmature_4p <- sites_cf %>%
  filter(plot_no == 4) %>%
  filter(mosaic > 0.5, mosaic < 0.76) %>%
  select(site) %>%
  left_join(as_tibble(sp_site, rownames = "site"), by = "site") 

community_secandmature_4p <- as.data.frame(community_secandmature_4p)
row.names(community_secandmature_4p) <- community_secandmature_4p$site
community_secandmature_4p <- community_secandmature_4p[,-1]


community_class2 <- sites_cf %>%
  filter(plot_no > 1) %>%
  filter(str_class == "2") %>%
  select(site) %>%
  left_join(as_tibble(sp_site, rownames = "site"), by = "site")

community_class2 <- as.data.frame(community_class2)
row.names(community_class2) <- community_class2$site
community_class2 <- community_class2[,-1]

community_class3 <- sites_cf %>%
  filter(plot_no > 1) %>%
  filter(str_class == "3") %>%
  select(site) %>%
  left_join(as_tibble(sp_site, rownames = "site"), by = "site")

community_class3 <- as.data.frame(community_class3)
row.names(community_class3) <- community_class3$site
community_class3 <- community_class3[,-1]

# Plots

community_class1_plots <- plots_cf %>%
  filter(str_class_plots == "1") %>%
  select(plot_id) %>%
  left_join(as_tibble(sp_site_plots, rownames = "plot_id"), by = "plot_id") 

community_class1_plots <- as.data.frame(community_class1_plots)
row.names(community_class1_plots) <- community_class1_plots$plot_id
community_class1_plots <- community_class1_plots[,-1]
  
community_class2_plots <- plots_cf %>%
  filter(str_class_plots == "2") %>%
  select(plot_id) %>%
  left_join(as_tibble(sp_site_plots, rownames = "plot_id"), by = "plot_id")

community_class2_plots <- as.data.frame(community_class2_plots)
row.names(community_class2_plots) <- community_class2_plots$plot_id
community_class2_plots <- community_class2_plots[,-1]

community_class3_plots <- plots_cf %>%
  filter(str_class_plots == "3") %>%
  select(plot_id) %>%
  left_join(as_tibble(sp_site_plots, rownames = "plot_id"), by = "plot_id")

community_class3_plots <- as.data.frame(community_class3_plots)
row.names(community_class3_plots) <- community_class3_plots$plot_id
community_class3_plots <- community_class3_plots[,-1]
```

Remove spp with abundance = 0 in each class

```{r}

# Sites

community_matureforest_4p <- community_matureforest_4p[, which(colSums(community_matureforest_4p) != 0)]
community_secandmature_4p <- community_secandmature_4p[, which(colSums(community_secandmature_4p) != 0)]
community_class2 <- community_class2[, which(colSums(community_class2) != 0)]
community_class3 <- community_class3[, which(colSums(community_class3) != 0)]

# Plots

community_class1_plots <- na.omit(community_class1_plots)
community_class1_plots <- community_class1_plots[, which(colSums(community_class1_plots) != 0)]

community_class2_plots <- na.omit(community_class2_plots)
community_class2_plots <- community_class2_plots[, which(colSums(community_class2_plots) != 0)]

community_class3_plots <- na.omit(community_class3_plots)
community_class3_plots <- community_class3_plots[,which(colSums(community_class3_plots) != 0)]

```

Use racurve function to plot rank-abundance curves

```{r}
racurve1 <- racurve(community_matureforest_4p, main = "Mature Forest", nlab = 8)
racurve2 <- racurve(community_secandmature_4p, main = "Secondary and Mature Forest", nlab = 8)
racurve2 <- racurve(community_class2, main = "Class 2", nlab = 5)
racurve3 <- racurve(community_class3, main = "Class 3", nlab = 5)

racurve(community_class2[1,], nlab = 5)
racurve(community_class2[2,], nlab = 5)
racurve(community_class2[3,], nlab = 5)
racurve(community_class2[4,], nlab = 5)
racurve(community_class2[5,], nlab = 5)
racurve(community_class2[6,], nlab = 5)
racurve(community_class2[7,], nlab = 5)

racurve(community_class3["70980",], nlab = 5)

# Plots
racurve1_plots <- racurve(community_class1_plots, main = "Class 1", nlab = 6)
racurve2_plots <- racurve(community_class2_plots, main = "Class 2", nlab = 6)
racurve3_plots <- racurve(community_class3_plots, main = "Class 3", nlab = 6)

```

```{r}
racurve3_plots
```

Relative abundance:

```{r}
racurve1$rel.abund
racurve2$rel.abund
racurve3$rel.abund

racurve1_plots$rel.abund
racurve2_plots$rel.abund
racurve3_plots$rel.abund
```

Let's take a closer look to each class
```{r}
sites_cf %>%
  filter(!is.na(str_class)) %>%
  filter(str_class == "1")

sites_cf %>%
  filter(!is.na(str_class)) %>%
  ggplot(aes(x = str_class, y = altitude)) +
  geom_point()

structure_forest <- sites_cf %>%
  filter(str_class == "2" | str_class == "3") %>%
  select(site,
         av_basalarea_site,
         av_loreys_site,
         av_treedensity_site)

# Scale data so that variables in different units can be compared
structure_forest_scaled <- scale(structure_forest[,-c(1)])
row.names(structure_forest_scaled) <- structure_forest$site

k2_structure_forest <- kmeans(structure_forest_scaled, centers = 2, nstart = 25)
k3_structure_forest <- kmeans(structure_forest_scaled, centers = 3, nstart = 25)
k4_structure_forest <- kmeans(structure_forest_scaled, centers = 4, nstart = 25)
k5_structure_forest <- kmeans(structure_forest_scaled, centers = 5, nstart = 25)

# Compare plots:
p1_str_f <- fviz_cluster(k2_structure_forest, geom = "point", data = structure_forest_scaled) + ggtitle("k = 2")
p2_str_f <- fviz_cluster(k3_structure_forest, geom = "point",  data = structure_forest_scaled) + ggtitle("k = 3")
p3_str_f <- fviz_cluster(k4_structure_forest, geom = "point",  data = structure_forest_scaled) + ggtitle("k = 4")
p4_str_f <- fviz_cluster(k5_structure_forest, geom = "point",  data = structure_forest_scaled) + ggtitle("k = 5")
grid.arrange(p1_str, p2_str, p3_str, p4_str, nrow = 2)


str_clusters_f <- NbClust((structure_forest_scaled^2), distance = "euclidean", min.nc = 2, max.nc = 10, method = "ward.D2", index = "all")

# Save a vector of 3 clusters:
str_k3_forest <- k3_structure_forest$cluster
as.numeric(str_k3_forest)

#pca with structural data

pca_str_f <- rda(structure_forest_scaled, scale = F)
summary(pca_str_f) 
plot(pca_str_f)

ordiplot(pca_str_f, display = 'sites', type = 'n')
points(pca_str_f, col = str_k3_forest)
biplot(pca_str_f, display = 'species')


```

Let's see if community data follows a similar classification than structural data

```{r}
# Calculate a distance matrix

log_sp_site <- log1p(sp_site)
dm_euclidean <- vegdist(log_sp_site, method = "euclidean")
cluster_ward <- hclust(dm_euclidean, method = "ward.D")
plot(cluster_ward)

library(dendextend)
dend_cluster_ward <- as.dendrogram(cluster_ward)
labels_colors(dend_cluster_ward)
labels_colors(dend_cluster_ward) <- as.numeric(sites_cf$str_class)
dend_cluster_ward <- set(dend_cluster_ward, "labels_cex", 0.5)
get_leaves_nodePar(dend_cluster_ward)[[1]]
plot(dend_cluster_ward)

```

