---
title: "05b_analysis"
author: "Adriana Uscanga"
date: "2023-02-07"
output: html_document
---

# Analysis: second round

Load packages and files

```{r}
# Files

load("output/sites_cf_4p.RData")
load("output/plots_cf_4p.RData")
load("output/trees_cf_4p.RData")

# Libraries

library(tidyverse)
library(ggplot2)
library(lme4)
library(corrplot)
library(vegan)
library(stargazer)
library(sjPlot)
library(cAIC4)

#library(rstatix)
#library(corrplot)
#library(vegan)
#library(leaps)
#library(asbio)
#library(spdep)
#library(ggpubr)
#library(multcompView)
```

## 1. Summary statistics and data exploration

```{r}

summary(plots_cf_4p)
summary(sites_cf_4p)

plots_cf_4p %>%
  ggplot(aes(x= agb_plot_ha)) +
  geom_histogram(bins = 25)

plots_cf_4p %>%
  ggplot(aes(x= log1p(agb_plot_ha))) +
  geom_histogram(bins = 20)

plots_cf_4p %>%
  ggplot(aes(x= sqrt(agb_plot_ha))) +
  geom_histogram(bins = 20)

plots_cf_4p %>%
  ggplot(aes(x= t_shannon)) +
  geom_histogram(bins = 20)

plots_cf_4p %>%
  ggplot(aes(x= sp_richness_plot)) +
  geom_histogram(bins = 10)

```

```{r}
numeric_variables <- plots_cf_4p %>%
  select(temp, precip, ai, altitude, latitude, longitude, northness, eastness, slope, t_shannon, sp_richness_plot, agb_plot_ha, agriculture, grazing, logging) %>%
  mutate(log_agb = log1p(agb_plot_ha)) %>%
  select(-agb_plot_ha)

corr <- cor(drop_na(numeric_variables[,-1]))
corrplot(corr, method = 'ellipse')
corrplot(corr, method = 'number')
plot(numeric_variables)
```
From this correlation matrix, I'm going to select variables that have correlation > 0.3
Environmental data is highly correlated, so I'm going to get rid of some of that collinearity with a PCA

```{r}
env <- plots_cf_4p %>%
  select(plot_id, temp, precip, ai, altitude, latitude, longitude, northness, eastness, slope)
env <- as.data.frame(env)
env$plot_id -> row.names(env)
env <- env[,c(-1, -2)]

pca_env <- rda(env, scale = T)
summary(pca_env)
plot(pca_env)
biplot(pca_env, display = 'species')

# PC1 (0.38) and PC2 (0.23) together explain 0.61 when lat and lon included
# PC1 (0.48) and PC2 (0.16) together explain 0.64 when lat and lon excluded

```

PC1:

```{r}
pc1 <- as.data.frame(pca_env$CA$u[,1])
colnames(pc1) <- "env_pc1"
pc1

plot(pc1[,1], plots_cf_4p$precip)
plot(pc1[,1], plots_cf_4p$ai)
plot(pc1[,1], plots_cf_4p$temp)
plot(pc1[,1], plots_cf_4p$altitude)

summary(lm(pc1[,1] ~ plots_cf_4p$precip))
summary(lm(pc1[,1] ~ plots_cf_4p$ai))
summary(lm(pc1[,1] ~ plots_cf_4p$temp))
summary(lm(pc1[,1] ~ plots_cf_4p$altitude))
```

PC2:

```{r}
pc2 <- as.data.frame(pca_env$CA$u[,2])
colnames(pc2) <- "env_pc2"
pc2

plot(pc2[,1], plots_cf_4p$slope)
plot(pc2[,1], plots_cf_4p$northness)
plot(pc2[,1], plots_cf_4p$eastness)
plot(pc2[,1], plots_cf_4p$latitude)
plot(pc2[,1], plots_cf_4p$longitude)

summary(lm(pc2[,1] ~ plots_cf_4p$slope))
summary(lm(pc2[,1] ~ plots_cf_4p$northness))
summary(lm(pc2[,1] ~ plots_cf_4p$eastness))
summary(lm(pc2[,1] ~ plots_cf_4p$latitude))
summary(lm(pc2[,1] ~ plots_cf_4p$longitude))

```
Slope and aspect are more related to PC2 but relationships are weak.
Correlation among them is also weak. 
PC2 is mainly driven by lat and lon.
To avoid collinearity, we are going to use PC1 (which encompasses temp, precip, elevation, and ai), and slope.
Lat, lon, and aspect don't seem to be related to other variables (see corrplot).

```{r}
# Add PC1 values to dataset

class(pc1)

env_pc1 <- as_tibble(pc1, .name_repair = "minimal", rownames = "plot_id")


plots_cf_4p <- plots_cf_4p %>%
  left_join(env_pc1, by = "plot_id")

plots_cf_4p <- plots_cf_4p %>%
  relocate(env_pc1, .after = ai)

```

Once we addressed collinearity, and checked which variables are correlated,
the selected variables that will go to the model are:
* env_pc1
* slope (?)
* t_shannon
* sp_richness_plot
* agriculture
* grazing
* logging (?)
* log_agb

(?) are not related to agb nor tree diversity but are correlated.

Mixed models

We are going to fit models for predicting AGB using the plots dataset
and using the variable 'site' as random factor.

1) First fit a model of agb with each predictor
2) then run a stepwise model selecion process
(backward and forward) with all predictors. 
3) Test interactions between predictors and assess whether adding
interactions in the model is worth it or not
4) repeat with glmm 


```{r}

lmm_env <- lmer(log1p(agb_plot_ha) ~ env_pc1 + (1| site), data = plots_cf_4p)
summary(lmm_env)
plot(lmm_env)
stargazer(lmm_env, type = "text")
cAIC(lmm_env)
```

```{r}
lmm_slope <- lmer(log1p(agb_plot_ha) ~ slope + (1| site), data = plots_cf_4p)
summary(lmm_slope)
plot(lmm_slope)
stargazer(lmm_slope, type = "text")
cAIC(lmm_slope)
```

```{r}

lmm_t_shannon <- lmer(log1p(agb_plot_ha) ~ t_shannon + (1| site), data = plots_cf_4p)
summary(lmm_t_shannon)
plot(lmm_t_shannon)
stargazer(lmm_t_shannon, type = "text")
cAIC(lmm_t_shannon)

```
```{r}

lmm_sp_rich <- lmer(log1p(agb_plot_ha) ~ sp_richness_plot + (1| site), data = plots_cf_4p)
summary(lmm_sp_rich)
plot(lmm_sp_rich)
stargazer(lmm_sp_rich, type = "text")
cAIC(lmm_sp_rich)

```
```{r}

lmm_ag <- lmer(log1p(agb_plot_ha) ~ agriculture + (1| site), data = plots_cf_4p)
summary(lmm_ag)
plot(lmm_ag)
stargazer(lmm_ag, type = "text")
cAIC(lmm_ag)

```
```{r}

lmm_graz <- lmer(log1p(agb_plot_ha) ~ grazing + (1| site), data = plots_cf_4p)
summary(lmm_graz)
plot(lmm_graz)
stargazer(lmm_graz, type = "text")
cAIC(lmm_graz)

```
```{r}

lmm_logg <- lmer(log1p(agb_plot_ha) ~ logging + (1| site), data = plots_cf_4p)
summary(lmm_logg)
plot(lmm_logg)
stargazer(lmm_logg, type = "text")
cAIC(lmm_logg)

```

```{r}
anocAIC(lmm_env, lmm_slope, lmm_t_shannon, lmm_sp_rich, lmm_ag, lmm_graz, lmm_logg, digits = 3)
```

Best model with only 1 predictor is agb ~ sp_richness!

```{r}

lmm_all <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           t_shannon +
           sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_all6 <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           t_shannon +
           sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_all5 <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           t_shannon +
           sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_all4 <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #t_shannon +
           sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)


anocAIC(lmm_all, lmm_all6, lmm_all5, lmm_all4, digits = 3)
```

"Stepwise" model (by hand)

```{r}
# Two predictors
combn(c("env","slope","sp_rich", "t_shannon","ag","graz","logg"), 2, simplify = T)

# env_pc1

lmm_2p_a <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #t_shannon +
           #sp_richness_plot +
           #agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_b <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           t_shannon +
           #sp_richness_plot +
           #agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_c <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           #t_shannon +
           sp_richness_plot +
           #agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_d <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           #t_shannon +
           #sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_e <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           #t_shannon +
           #sp_richness_plot +
           #agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_f <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           #t_shannon +
           #sp_richness_plot +
           #agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

# Slope 

lmm_2p_g <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           t_shannon +
           #sp_richness_plot +
           #agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_h <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           #t_shannon +
           sp_richness_plot +
           #agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_i <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           #t_shannon +
           #sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_j <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           #t_shannon +
           #sp_richness_plot +
           #agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_k <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           #t_shannon +
           #sp_richness_plot +
           #agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

# T_shannon

lmm_2p_l <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           t_shannon +
           sp_richness_plot +
           #agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_m <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           t_shannon +
           #sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_n <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           t_shannon +
           #sp_richness_plot +
           #agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_o <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           t_shannon +
           #sp_richness_plot +
           #agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

# Sp_richness_plot

lmm_2p_p <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           #t_shannon +
           sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_q <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           #t_shannon +
           sp_richness_plot +
           #agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_r <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           #t_shannon +
           sp_richness_plot +
           #agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

# Agriculture

lmm_2p_s <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           #t_shannon +
           #sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_2p_t <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           #t_shannon +
           #sp_richness_plot +
           agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

# Grazing

lmm_2p_u <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           #t_shannon +
           #sp_richness_plot +
           #agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm2p_cAIC <- anocAIC(lmm_2p_a,
        lmm_2p_b,
        lmm_2p_c,
        lmm_2p_d,
        lmm_2p_e,
        lmm_2p_f,
        lmm_2p_g,
        lmm_2p_h,
        lmm_2p_i,
        lmm_2p_j,
        lmm_2p_k,
        lmm_2p_l,
        lmm_2p_m,
        lmm_2p_n,
        lmm_2p_o,
        lmm_2p_p,
        lmm_2p_q,
        lmm_2p_r,
        lmm_2p_s,
        lmm_2p_t,
        lmm_2p_u,
        digits = 3)

lmm2p_cAIC <- as_tibble(lmm2p_cAIC, rownames = "model")

lmm2p_cAIC %>%
  arrange(cAIC)
```
Best model with 2 predictors: sp_richness + agriculture cAIC 446.844
I'll remove t_shannon because sp_richness seems to work better and they're very related

```{r}
# 3 predictors
combn(c("env","slope","sp_rich","ag","graz","logg"), 3, simplify = T)

# env_pc1

lmm_3p_a <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           sp_richness_plot +
           #agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_b <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_c <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #sp_richness_plot +
           #agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_d <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #sp_richness_plot +
           #agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_3p_e <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_f <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           sp_richness_plot +
           #agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_g <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           sp_richness_plot +
           #agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_3p_h <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           #sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_i <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           #sp_richness_plot +
           agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_3p_j <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           #sp_richness_plot +
           #agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

# slope

lmm_3p_k <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_l <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           sp_richness_plot +
           #agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_m <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           sp_richness_plot +
           #agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_3p_n <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           #sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_o <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           #sp_richness_plot +
           agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_3p_p <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           #sp_richness_plot +
           #agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

# sp_richness

lmm_3p_q <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_3p_r <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           sp_richness_plot +
           agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_3p_s <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           sp_richness_plot +
           #agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

# agriculture

lmm_3p_t <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           #sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm3p_cAIC <- anocAIC(lmm_3p_a,
        lmm_3p_b,
        lmm_3p_c,
        lmm_3p_d,
        lmm_3p_e,
        lmm_3p_f,
        lmm_3p_g,
        lmm_3p_h,
        lmm_3p_i,
        lmm_3p_j,
        lmm_3p_k,
        lmm_3p_l,
        lmm_3p_m,
        lmm_3p_n,
        lmm_3p_o,
        lmm_3p_p,
        lmm_3p_q,
        lmm_3p_r,
        lmm_3p_s,
        lmm_3p_t,
        digits = 3)

lmm3p_cAIC <- as_tibble(lmm3p_cAIC, rownames = "model")

lmm3p_cAIC %>%
  arrange(cAIC)

```
Best model with 3 predictors is slope + sp_richness + agriculture cAIC 443.707

```{r}
# 4 predictors
combn(c("env","slope","sp_rich","ag","graz","logg"), 4, simplify = T)

# env_pc1

lmm_4p_a <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           sp_richness_plot +
           agriculture +
           #grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_4p_b <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           sp_richness_plot +
           #agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_4p_c <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           sp_richness_plot +
           #agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_d <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_4p_e <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #sp_richness_plot +
           agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_f <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #sp_richness_plot +
           #agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_g <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_4p_h <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           sp_richness_plot +
           agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_i <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           sp_richness_plot +
           #agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_j <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           #sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_k <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_4p_l <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           sp_richness_plot +
           agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_m <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           sp_richness_plot +
           #agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_n <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           #sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_4p_o <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           #slope +
           sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm4p_cAIC <- anocAIC(lmm_4p_a,
        lmm_4p_b,
        lmm_4p_c,
        lmm_4p_d,
        lmm_4p_e,
        lmm_4p_f,
        lmm_4p_g,
        lmm_4p_h,
        lmm_4p_i,
        lmm_4p_j,
        lmm_4p_k,
        lmm_4p_l,
        lmm_4p_m,
        lmm_4p_n,
        lmm_4p_o,
        digits = 3)

lmm4p_cAIC <- as_tibble(lmm4p_cAIC, rownames = "model")

lmm4p_cAIC %>%
  arrange(cAIC)

```
Best model with 4 predictors is env_pc1 + slope + sp_richness + agriculture cAIC: 443.194

```{r}

# 5 predictors
combn(c("env","slope","sp_rich","ag","graz","logg"), 5, simplify = T)

lmm_5p_a <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           sp_richness_plot +
           agriculture +
           grazing +
           #logging +
           (1| site), data = plots_cf_4p)

lmm_5p_b <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           sp_richness_plot +
           agriculture +
           #grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_5p_c <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           sp_richness_plot +
           #agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_5p_d <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           #sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_5p_e <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           #slope +
           sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm_5p_f <- lmer(log1p(agb_plot_ha) ~ #env_pc1 +
           slope +
           sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

lmm5p_cAIC <- anocAIC(lmm_5p_a,
        lmm_5p_b,
        lmm_5p_c,
        lmm_5p_d,
        lmm_5p_e,
        lmm_5p_f,
        digits = 3)

lmm5p_cAIC <- as_tibble(lmm5p_cAIC, rownames = "model")

lmm5p_cAIC %>%
  arrange(cAIC)

```

Best model with 5 predictors is env_pc1 + slope + sp_richness_plot + agriculture + logging cAIC 443.769

```{r}
# All 6 predictors

lmm_6p_a <- lmer(log1p(agb_plot_ha) ~ env_pc1 +
           slope +
           sp_richness_plot +
           agriculture +
           grazing +
           logging +
           (1| site), data = plots_cf_4p)

cAIC(lmm_6p_a)
```

Best model has 4 predictors: 
```{r}

summary(lmm_4p_a)
plot(lmm_4p_a)
plot_model(lmm_4p_a, show.p = T, show.values = T)
stargazer(lmm_4p_a, type = "text")
cAIC(lmm_4p_a)

qqnorm(resid(lmm_4p_a))
qqline(resid(lmm_4p_a))

shapiro.test(resid(lmm_4p_a))
```

As we would expect some variables to be related, we also made models with interactions

```{r}


lmm_4p_int <- lmer(log1p(agb_plot_ha) ~ env_pc1 * agriculture +
                     sp_richness_plot * agriculture +
                     slope +
                     (1| site), data = plots_cf_4p)

cAIC(lmm_4p_int)
stargazer(lmm_4p_int, type = "text")

summary(lmm_4p_int)
shapiro.test(resid(lmm_4p_int))

plot(lmm_4p_int)
plot_model(lmm_4p_int, show.p = T, show.values = T)

qqnorm(resid(lmm_4p_int))
qqline(resid(lmm_4p_int))


```

```{r}
cAIC(lmm_4p_a)
cAIC(lmm_4p_int)
```

So far, the best models are with 4 predictors. 
When adding an interactions between agriculture and sp_richness, the model improves a lot,
although residuals are not normal.



Linear mixed model

LMM with all variables and site as random effect

```{r}
lmm <- lmer(log1p(agb_plot_ha) ~ env_pc1 + slope + northness + eastness + t_shannon + agriculture + grazing + (1| site), data = plots_cf_4p)

summary(lmm)
plot(lmm)
qqnorm(resid(lmm))
qqline(resid(lmm))

plot_model(lmm, show.values = T)
stargazer(lmm, type = "text")
```

LMM with all variables, interactions among them, and sites as random effect

```{r}
lmm2 <- lmer(log1p(agb_plot_ha) ~ env_pc1 + slope + northness + eastness + t_shannon + agriculture + grazing + env_pc1*slope + slope*agriculture + slope*t_shannon + slope*grazing + env_pc1*grazing + env_pc1*agriculture + env_pc1*t_shannon + t_shannon*grazing + t_shannon*agriculture + grazing*agriculture + (1|site), data = plots_cf_4p)

summary(lmm2)
plot(lmm2)
qqnorm(resid(lmm2))
qqline(resid(lmm2))

plot_model(lmm2, show.values = T)

stargazer(lmm2, type = "text")
```
Lower AIC. 
I'll remove non-relevant variables first, and then non-relevant interactions:

```{r}
lmm3 <- lmer(log1p(agb_plot_ha) ~ env_pc1 + slope + t_shannon + agriculture + grazing + (1|site), data = plots_cf_4p)

summary(lmm3)
plot(lmm3)
qqnorm(resid(lmm3))
qqline(resid(lmm3))

plot_model(lmm3, show.values = T)

stargazer(lmm3, type = "text")
```
This is the best model so far, although grazing is non-significant.
Let's add the relevant interactions:

```{r}
lmm4 <- lmer(log1p(agb_plot_ha) ~ env_pc1 + slope + t_shannon + agriculture + t_shannon*agriculture + slope*t_shannon + env_pc1*grazing + (1|site), data = plots_cf_4p)

summary(lmm4)
stargazer(lmm4, type = "text")
plot_model(lmm4, show.values = T)

```
I'll remove the only interaction that wasn't statistically significant:

```{r}
lmm5 <- lmer(log1p(agb_plot_ha) ~ env_pc1 + slope + t_shannon + agriculture + t_shannon*agriculture + env_pc1*grazing + (1|site), data = plots_cf_4p)

summary(lmm5)
stargazer(lmm5, type = "text")
plot_model(lmm5, show.values = T)
```
Best model thus far.

```{r}

plots_cf_4p %>%
  ggplot(aes(x = t_shannon, y = log1p(agb_plot_ha))) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(y = t_shannon, x = agriculture)) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(y = t_shannon, x = grazing)) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = agriculture, y = log1p(agb_plot_ha))) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = slope, y = log1p(agb_plot_ha))) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = grazing, y = log1p(agb_plot_ha))) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = env_pc1, y = agriculture)) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = altitude, y = agriculture)) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = altitude, y = grazing)) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = env_pc1, y = grazing)) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = env_pc1, y = t_shannon)) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

plots_cf_4p %>%
  ggplot(aes(x = env_pc1, y = log1p(agb_plot_ha))) +
  geom_point(aes(color = str_class_plots)) +
  geom_smooth(method = "lm")

```

GLMM: agb ~ diversity

Name | Link Function | Distribution
General linear model | Identity | Normal
Logistic regression | Logit | Binomial
Poisson | Log | Poisson
Gamma | Inverse | Gamma

```{r}

glmm1 <- glmer((scale(agb_plot_ha)+1) ~ t_shannon + (1|site), Gamma(link = "inverse"), data = plots_cf_4p, nAGQ = 2, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm1)

glmm1b <- glmer((scale(agb_plot_ha)+1) ~ sp_richness_plot + (1|site), Gamma(link = "inverse"), data = plots_cf_4p, nAGQ = 2, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm1b)

lmer1 <- lmer(log1p(agb_plot_ha) ~ t_shannon + (1|site), data = plots_cf_4p)
summary(lmer1)
plot(lmer1)
plot_model(lmer1, show.p = T, show.values = T)
stargazer(lmer1, type = "text")
# This model is significant but the relationship doesn't seem to be normal
# Also, about 60% of variation is explained by the random effect
# which I think it means that there's a stronger relationship between
# agb and diversity between plots than between sites 

lmer1b <- lmer(agb_plot_ha ~ t_shannon + (1|site), data = plots_cf_4p)
summary(lmer1b)
plot(lmer1b)
stargazer(lmer1b, type = "text")

glmm1c <- glmer((log1p(agb_plot_ha+1)) ~ t_shannon + (1|site), Gamma(link = "inverse"), data = plots_cf_4p, nAGQ = 2, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm1c)
plot(glmm1c)

# GAMM?
```

GMM: agb ~ all variables

```{r}

glmm2 <- glmer((scale(agb_plot_ha)+1) ~ env_pc1 + scale(slope) + t_shannon + agriculture + grazing + env_pc1*scale(slope) + scale(slope)*agriculture + scale(slope)*t_shannon + scale(slope)*grazing + env_pc1*grazing + env_pc1*agriculture + env_pc1*t_shannon + t_shannon*grazing + t_shannon*agriculture + grazing*agriculture + (1|site), Gamma(link="inverse"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm2)

# dropping variables that are not significant in 'all variables and all interactions model'
glmm2b <- glmer((scale(agb_plot_ha)+1) ~ env_pc1 + scale(slope) + agriculture + scale(slope)*agriculture + scale(slope)*grazing + env_pc1*grazing + t_shannon*agriculture + grazing*agriculture + (1|site), Gamma(link="inverse"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm2b)

# Agriculture ~ environmental gradient
g <- glmer.nb(agriculture ~ env_pc1 + (1|site), data = plots_cf_4p, verbose = T)
getME(g, "glmer.nb.theta")

glmm3 <- glmer(agriculture ~ env_pc1 + (1|site), negative.binomial(theta= 87730.47, link="log"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm3)

## *Significant interaction!*

# Agriculture ~ slope

gb <- glmer.nb(agriculture ~ scale(slope) + (1|site), data = plots_cf_4p, verbose = T)
getME(gb, "glmer.nb.theta")

glmm3b <- glmer(agriculture ~ scale(slope) + (1|site), negative.binomial(theta= 88474.5, link="log"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm3b)

## Non-significant

# Grazing ~ environmental gradient

# Grazing only has two values so I'm making it binomial 
ds <- plots_cf_4p %>%
  mutate(graz_bi = ifelse(grazing == 2, 1, grazing)) 
glmm4 <- glmer(graz_bi ~ env_pc1 + (1|site), binomial(link="logit"), data = ds, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm4) #as binomial variable

g2 <- glmer.nb(grazing ~ env_pc1 + (1|site), data = plots_cf_4p, verbose = T)
getME(g2, "glmer.nb.theta")

glmm4b <- glmer(grazing ~ env_pc1 + (1|site), negative.binomial(theta= 75716.24, link="log"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm4b) #as discrete variable

## Non-significant

# Grazing ~ slope

glmm4c <- glmer(graz_bi ~ scale(slope) + (1|site), binomial(link="logit"), data = ds, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm4c) #as binomial variable

## Non-significant

# Tree diversity (t_shannon) ~ environmental gradient

lmer_tshannon1 <- lmer(sqrt(t_shannon) ~ env_pc1 + (1|site), data = plots_cf_4p)
summary(lmer_tshannon1)
stargazer(lmer_tshannon1, type = "text")
plot_model(lmer_tshannon1, show.values = T)
qqnorm(resid(lmer_tshannon1))
qqline(resid(lmer_tshannon1))

## Non-significant

# Tree diversity ~ agriculture

glmm5b <- glmer((t_shannon+0.1) ~ agriculture + (1|site), Gamma(link="inverse"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm5b)

lmer_tshannon <- lmer(t_shannon ~ agriculture + (1|site), data = plots_cf_4p)
summary(lmer_tshannon)
stargazer(lmer_tshannon, type = "text")
plot_model(lmer_tshannon, show.values = T)
qqnorm(resid(lmer_tshannon))
qqline(resid(lmer_tshannon))

# sqrt transformation for slightly right-skewed data
lmer_tshannon2 <- lmer(sqrt(t_shannon) ~ agriculture + (1|site), data = plots_cf_4p)
summary(lmer_tshannon2)
stargazer(lmer_tshannon2, type = "text")
plot_model(lmer_tshannon2, show.values = T)
qqnorm(resid(lmer_tshannon2))
qqline(resid(lmer_tshannon2))

## *Significant!*

# Tree diversity ~ grazing

lmer_tshannon3 <- lmer(sqrt(t_shannon) ~ grazing + (1|site), data = plots_cf_4p)
summary(lmer_tshannon3)
stargazer(lmer_tshannon3, type = "text")
plot_model(lmer_tshannon3, show.values = T)
qqnorm(resid(lmer_tshannon3))
qqline(resid(lmer_tshannon3))

## *Significant!*

# Tree diversity ~ slope 
lmer_tshannon_slope <- lmer(sqrt(t_shannon) ~ scale(slope) + (1|site), data = plots_cf_4p)
summary(lmer_tshannon_slope)
stargazer(lmer_tshannon_slope, type = "text")

## Non-significant
```


With interactions:

```{r}

glmm_w_interactions <- glmer((scale(agb_plot_ha)+1) ~  env_pc1*agriculture + 
                                         scale(slope) +
                                         t_shannon*agriculture + 
                                         t_shannon*grazing +
                                         (1|site), Gamma(link="inverse"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm_w_interactions)
plot(glmm_w_interactions)
qqnorm(resid(glmm_w_interactions))
qqline(resid(glmm_w_interactions))

plot_model(glmm_w_interactions, show.values = T)
stargazer(glmm_w_interactions, type = "text")

# Reduced:

glmm_w_interactions_r <- glmer((scale(agb_plot_ha)+1) ~  env_pc1*agriculture + 
                                         scale(slope) +
                                         t_shannon*agriculture + 
                                         #t_shannon*grazing +
                                         (1|site), Gamma(link="inverse"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm_w_interactions_r)
plot(glmm_w_interactions_r)
qqnorm(resid(glmm_w_interactions_r))
qqline(resid(glmm_w_interactions_r))

plot_model(glmm_w_interactions_r, show.values = T)
stargazer(glmm_w_interactions_r, type = "text")

glmm_wo_interactions <- glmer((scale(agb_plot_ha)+1) ~  env_pc1 + 
                                                        agriculture + 
                                                        scale(slope) +
                                                        t_shannon + 
                                                        grazing +
                                         (1|site), Gamma(link="inverse"), data = plots_cf_4p, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
summary(glmm_wo_interactions)
plot(glmm_wo_interactions)
qqnorm(resid(glmm_wo_interactions))
qqline(resid(glmm_wo_interactions))

plot_model(glmm_wo_interactions, show.values = T)
stargazer(glmm_wo_interactions, type = "text")
```

```{r}

lmer_w_interactions <- lmer(log1p(agb_plot_ha) ~ env_pc1*agriculture + 
                                         #env_pc1*grazing + #If added to the model, AIC decreases and env turns very relevant, but it seems weird because these two are not really related (see test above)
                                         scale(slope) +
                                         t_shannon*agriculture + 
                                         t_shannon*grazing +
                                         (1|site), data = plots_cf_4p)

summary(lmer_w_interactions)
plot(lmer_w_interactions)
qqnorm(resid(lmer_w_interactions))
qqline(resid(lmer_w_interactions))

plot_model(lmer_w_interactions, show.values = T)
stargazer(lmer_w_interactions, type = "text")
```


Plots classification

Option 1: k-means including PC1
```{r}
plots_str2 <- plots_cf_4p %>%
  select(plot_id,
         tree_density,
         loreys_height,
         basal_area_ha,
         agb_plot_ha,
         env_pc1)

plot(plots_str2)

# Scale data so that variables in different units can be compared
plots_str_scaled2 <- scale(plots_str2[,-c(1,2,5)])
row.names(plots_str_scaled2) <- plots_str2$plot_id

k2_structure_plots2 <- kmeans(plots_str_scaled2, centers = 2, nstart = 25)
k3_structure_plots2 <- kmeans(plots_str_scaled2, centers = 3, nstart = 25)
k4_structure_plots2 <- kmeans(plots_str_scaled2, centers = 4, nstart = 25)
k5_structure_plots2 <- kmeans(plots_str_scaled2, centers = 5, nstart = 25)

p1_str_plots2 <- fviz_cluster(k2_structure_plots2, geom = "point", data = plots_str_scaled2) + ggtitle("k = 2")
p2_str_plots2 <- fviz_cluster(k3_structure_plots2, geom = "point",  data = plots_str_scaled2) + ggtitle("k = 3")
p3_str_plots2 <- fviz_cluster(k4_structure_plots2, geom = "point",  data = plots_str_scaled2) + ggtitle("k = 4")
p4_str_plots2 <- fviz_cluster(k5_structure_plots2, geom = "point",  data = plots_str_scaled2) + ggtitle("k = 5")
grid.arrange(p1_str_plots2, p2_str_plots2, p3_str_plots2, p4_str_plots2, nrow = 2)

str_clusters_plots2 <- NbClust((plots_str_scaled2^2), distance = "euclidean", min.nc = 2, max.nc = 10, method = "ward.D2", index = "all")

# Save a vector of 2 clusters:
str_k2_plots2 <- k2_structure_plots2$cluster
as.numeric(str_k2_plots2)

#pca with structural data

pca_str_plots2 <- rda(plots_str_scaled2, scale = F)
summary(pca_str_plots2) # 
plot(pca_str_plots2)

ordiplot(pca_str_plots2, display = 'sites', type = 'n')
points(pca_str_plots2, col = str_k2_plots2)
biplot(pca_str_plots2, display = 'species')

ordiplot(pca_str_plots, display = 'sites', type = 'text')
biplot(pca_str_plots)

ordiplot(pca_str_plots2, display = 'sites', type = 'n')
points(pca_str_plots, col = str_k3_plots)
biplot(pca_str_plots, display = 'species')

ordiplot(pca_str_plots, display = 'sites', type = 'text')
biplot(pca_str_plots)

plots_cf_4p %>%
  ggplot(aes(x= str_class_plots, y = agriculture, color = basal_area_ha)) +
  geom_jitter()

plots_cf_4p %>%
  ggplot(aes(x= str_class_plots, y = grazing)) +
  geom_jitter(color = 'blue')

plots_cf_4p %>%
  filter(str_class_plots == 1) %>%
  ggplot(aes(x = env_pc1, y = loreys_height, color = as.factor(agriculture))) +
  geom_point()

plots_cf_4p %>%
  filter(str_class_plots == 2) %>%
  ggplot(aes(x = altitude, y = agb_plot_ha, color = as.factor(agriculture))) +
  geom_point()

plots_cf_4p %>%
  filter(str_class_plots == 3) %>%
  ggplot(aes(x = altitude, y = agb_plot_ha, color = as.factor(agriculture))) +
  geom_point()


plots_cf_4p %>%
  filter(str_class_plots == 1) %>%
  arrange(agriculture, -altitude)

plots_cf_4p %>%
  mutate(land_use = agriculture + grazing) %>%
  ggplot(aes(x = str_class_plots, y = land_use)) +
  geom_jitter()
  
```


Option 1: PCA?